Validation of Visual Statistical Inference, Applied to Linear Models

### 2012-06-28 paper_jasa.tex
added a model fit section

Let $X_{\ell i} \sim Binom_{1, p_{\ell i}}$ be the random variable, with $X_{ \ell i}= 1$, if  subject $i$ correctly identifies the data plot from lineup $\ell$, and 0 otherwise.
We can model $p_{ \ell i}$ with a logistic regression with subject specific random effects as:
\begin{equation} \label{mixed}
g( p_{\ell i} )= W_{\ell i} \delta +  Z_{\ell i}  \tau_{\ell i},  
\end{equation}

@@

@@

It can also be seen that at effect $E=0$, the power is close to 0.05 (Type-I error) for both experiments 1 and 2. For experiment 3 the scenario is different. The contamination in the data is visible which sets the lower limit of proportion correct at a higher value. For this reason, when the subject-specific models (Equation \ref{mixed}) are fit to the data, in the next section, we set offset = $log(0.05/0.95)$ so that the estimated power has a fixed lower limit at 0.05 (Type-I error) when $E=0$. While fitting models (Equation \ref{mixed}) we set $W_{\ell i} = E_{\ell i}$, the effect of lineup $\ell$ that subject $i$ has evaluated. Also $E$ is used as the only fixed effect covariate without intercept but offset for experiment 1 and 2. For experiment 3 an intercept is estimated to allow the power curves to go above 0.05.

@@

\green{Do we have more than one model?? If so then it is appropriate to refer to them as model 1, model 2, ...., but if there is only one model then we need to refer readers back to the equation, occasionally.}

\green{Actually I think this paragraph should have already been covered when the model was introduced. At this point we just need to discuss the results, so I'm going to cut the paragraph and start again.}

%Different subjects are expected to have different skill levels in evaluating lineups. This is accounted for in the mixed effects model (Equation \ref{mixed}).  For all the three experiments, each subject saw a lineup just once. This doesn't allow fitting a lineup random effect, in addition to a subject random effect, but it does simplify the model since $\tau_{\ell i}=\tau_i$. The subject-specific probability ($p_i$) of successful evaluation of a lineup can be estimated in the model as $\hat p_i= ....$. So, we include a random interaction term with effect. 
Table \ref{tbl:model_par} shows the parameter estimates of the mixed effects model of the subject-specific variation.  The estimated subject specific power curves are shown in Figure \ref{fig:power_mixed_subject} (light grey lines). The blue curve is the overall average power estimated from the model, and for comparison, the grey dashed lines are the power curves for the conventional test. Subject-specific power is quite different for three experiments. In experiment 1 subjects performed similarly, and substantially worse than the conventional test. In experiment 2 there is more variability between subjects, with some doing better than the conventional test on small effects. In experiment 3 there is the most subject-specific variation. Some subjects performed substantially better than the conventional test, and on average the visual test was better.

@@

\begin{figure}[hbtp]
   \centering
       \scalebox{0.60}{\includegraphics{power_mixed_subject.pdf}}
       \caption{Subject specific  power estimated from model \ref{mixed} are shown. The corresponding power curve for conventional test (dashed line) is shown for comparison. The overall average power curve estimated from the model is shown (light blue).}
       \label{fig:power_mixed_subject}
\end{figure}

\subsection{Estimating $p$-value in non-classical settings}

@@

\begin{figure}[hbtp]
   \centering
       \scalebox{0.60}{\includegraphics{power_screening_subject.pdf}}
      \caption{The subject specific power estimated from model \ref{mixed} are shown for various data sets obtained by applying 6 screening criteria.  The corresponding power curve for conventional test is shown in pink lines for comparison.}
       \label{fig:power_screening_subject}	
\end{figure}

>>>>

Let $X_{\ell i} \sim Binom_{1, p_{\ell i}}$ be the random variable, with $X_{ \ell i}= 1$, if  subject $i$ correctly identifies the data plot from lineup $\ell$, and 0 otherwise.
We can model $p_{ \ell i}$ with a logistic regression with subject-specific random effects as:
\begin{equation} \label{mixed}
g( p_{\ell i} )= W_{\ell i} \delta +  Z_{\ell i}  \tau_{\ell i},  
\end{equation}

@@

\subsection{Model fitting}

The model in Equation \ref{mixed}is fit using \green{.... fill this in} Table \ref{tbl:model_par} shows the parameter estimates of the mixed effects model of the subject-specific variation. For all the three experiments, each subject saw a lineup just once. This doesn't allow fitting a lineup random effect, in addition to a subject random effect, but it does simplify the model since $\tau_{\ell
  i}=\tau_i$. The subject-specific probability ($p_i$) of successful evaluation of a lineup can be estimated in the model as $\hat p_i= ....$. So, we include a random interaction term with effect.

Different subjects are expected to have different skill levels in evaluating lineups. For this reason, when the subject-specific models (Equation \ref{mixed}) are fit to the data, in the next section, we set offset = $log(0.05/0.95)$ so that the estimated power has a fixed lower limit at 0.05 (Type-I error) when $E=0$. While fitting models (Equation \ref{mixed}) we set $W_{\ell i} = E_{\ell i}$, the effect of lineup $\ell$ that subject $i$ has evaluated. Also $E$ is used as the only fixed effect covariate without intercept but offset for experiment 1 and 2. For experiment 3 an intercept is estimated to allow the power curves to go above 0.05.

\green{This section needs to have sufficient detail of the model fitting, output and diagnostics to feed into the next few sections discussing the findings.}

@@

It can also be seen that at effect $E=0$, the power is close to 0.05 (Type-I error) for both experiments 1 and 2. For experiment 3 the scenario is different. The contamination in the data is visible which sets the lower limit of proportion correct at a higher value.

@@

%Different subjects are expected to have different skill levels in evaluating lineups. This is accounted for in the mixed effects model (Equation \ref{mixed}).   
 The estimated subject-specific power curves are shown in Figure \ref{fig:power_mixed_subject} (light grey lines). The blue curve is the overall average power estimated from the model, and for comparison, the grey dashed lines are the power curves for the conventional test. Subject-specific power is quite different for three experiments. In experiment 1 subjects performed similarly, and substantially worse than the conventional test. In experiment 2 there is more variability between subjects, with some doing better than the conventional test on small effects. In experiment 3 there is the most subject-specific variation. Some subjects performed substantially better than the conventional test, and on average the visual test was better.

@@

\begin{figure}[hbtp]
   \centering
       \scalebox{0.60}{\includegraphics{power_mixed_subject.pdf}}
       \caption{Subject-specific  power estimated from model \ref{mixed} are shown. The corresponding power curve for conventional test (dashed line) is shown for comparison. The overall average power curve estimated from the model is shown (light blue).}
       \label{fig:power_mixed_subject}
\end{figure}

\subsection{Estimating the $p$-value in the real world}

@@

\begin{figure}[hbtp]
   \centering
       \scalebox{0.60}{\includegraphics{power_screening_subject.pdf}}
      \caption{The subject-specific power estimated from model \ref{mixed} are shown for various data sets obtained by applying 6 screening criteria.  The corresponding power curve for conventional test is shown in pink lines for comparison.}
       \label{fig:power_screening_subject}	
\end{figure}
